---
title: "Treasurer's Report"
author: "David Barron"
date: "`r format(Sys.Date(), '%d %B %Y' )`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)

opts_chunk$set(echo = FALSE)
options(knitr.kable.NA = '')

# Format money output
inline_hook <- function(x){
  format(x, nsmall = 2, scientific = FALSE, big.mark = ",")
}

knit_hooks$set(inline = inline_hook)

```

```{r intro, message=FALSE}
library(readxl)
library(tidyverse)

year <- 2025  # for filename
prevYear <- 2024

```


```{r data, message=FALSE}

fn <- paste0("C:\\Users\\dnbar\\Dropbox\\OAS\\OAS_Finances_", year, "\\OAS ", year, ".xlsx")

read_dta <- function(fn, sheet){
  read_xlsx(fn, sheet = sheet) %>%
    filter(Category != "Transfer Out") %>%  # remove transfers between accounts
    filter(Category != "Transfer In")
}

dta <- read_dta(fn, "Current account")
savings.dta <- read_dta(fn, "Savings account")

nrows <- nrow(dta)

# Summary of income and expenditure by month

sum_dta <- function(dta){
  dta %>%
    mutate(Date = lubridate::ymd(Date),
           Month = lubridate::month(Date)) %>%
      group_by(Month) %>%
      summarise(
          Expend = sum(Expenditure, na.rm = TRUE),
          Inc = sum(Income, na.rm = TRUE)
      )  
}

dta.sum <- sum_dta(dta)
savings.dta.sum <- sum_dta(savings.dta)

# calculate cummulative amounts for plotting
calc.cumsum <- function(dta){
  dta %>%
    mutate(
          Income = cumsum(Inc),
          Expenditure = cumsum(Expend)
    ) %>%
    select(Month, Income, Expenditure)
}

dta.cumsum <- calc.cumsum(dta.sum)
savings.dta.cumsum <- calc.cumsum(savings.dta.sum)
  
num.months <- dim(dta.cumsum)[1]
num.months.sav <- dim(savings.dta.cumsum)[1]

# Transform data for plotting

calc.long <- function(dta) {
  tmp <- dta %>% pivot_longer(
              cols = Income:Expenditure,
              names_to = "Type",
              values_to = "Amount"
  ) 
}

dta.long <- calc.long(dta.cumsum)
savings.dta.long <- calc.long(savings.dta.cumsum)


# Add together amounts from current and savings accounts
# Have to take account of times when there have been no changes to savings account
#  during the current month

total.dta.long.dta <- full_join(dta.long, savings.dta.long, 
                            by = c("Month", "Type"))
total.dta.long.dta[is.na(total.dta.long.dta[, 4]), 4] <- 0

total.dta.long <- total.dta.long.dta %>%
  mutate(Amount = Amount.x + Amount.y,
         Year = year) %>%  #need a year variable for when data is combined
  select(Month, Type, Amount, Year)


nrow.savings <- nrow(savings.dta)
savings.balance <- as.numeric(savings.dta[nrow.savings, 6])

calc.sum <- function(dta){
  dta %>%
  group_by(Category) %>%
  summarise(Income = sum(Income, na.rm = TRUE),
            Expenditure = sum(Expenditure, na.rm = TRUE))
}

sum_tab <- calc.sum(dta)

savings.sum_tab <- calc.sum(savings.dta)
  
total.income <- sum(sum_tab$Income) 

total.expenditure <- sum(sum_tab$Expenditure) 


```


```{r prevdata, message=FALSE, warning=FALSE}

# Read and process previous year's data
prevfn <- paste0("C:\\Users\\dnbar\\Dropbox\\OAS\\", prevYear, "\\OAS ", prevYear, ".xlsx")


prev.dta <- read_dta(prevfn, "Current account")
prev.savings.dta <- read_dta(prevfn, "Savings account")

prev.nrows <- nrow(prev.dta)

# Summary of income and expenditure by month

prev.dta.sum <- sum_dta(prev.dta)
prev.savings.dta.sum <- sum_dta(prev.savings.dta)

# calculate cummulative amounts for plotting

prev.dta.cumsum <- calc.cumsum(prev.dta.sum)
prev.savings.dta.cumsum <- calc.cumsum(prev.savings.dta.sum)
  
prev.num.months <- dim(prev.dta.cumsum)[1]
prev.num.months.sav <- dim(prev.savings.dta.cumsum)[1]

# Transform data for plotting


prev.dta.long <- calc.long(prev.dta.cumsum)
prev.savings.dta.long <- calc.long(prev.savings.dta.cumsum)


# Add together amounts from current and savings accounts
# Have to take account of times when there have been no changes to savings account
#  during the current month

prev.total.dta.long.dta <- full_join(prev.dta.long, prev.savings.dta.long, 
                            by = c("Month", "Type"))
prev.total.dta.long.dta[is.na(prev.total.dta.long.dta[, 4]), 4] <- 0

prev.total.dta.long <- prev.total.dta.long.dta %>%
  mutate(Amount = Amount.x + Amount.y,
         Year = prevYear) %>%  # need a Year variable for when data is combined
  filter(!is.na(Month)) %>%
  select(Month, Type, Amount, Year)


prev.nrow.savings <- nrow(prev.savings.dta)
prev.savings.balance <- as.numeric(prev.savings.dta[prev.nrow.savings, 6])

prev.sum_tab <- calc.sum(prev.dta)

prev.savings.sum_tab <- calc.sum(prev.savings.dta)
  
prev.total.income <- as.numeric(prev.dta.cumsum[prev.num.months, 2]) + as.numeric(prev.savings.dta.cumsum[prev.num.months.sav, 2]) 

prev.total.expenditure <- as.numeric(prev.dta.cumsum[prev.num.months, 3]) + as.numeric(prev.savings.dta.cumsum[prev.num.months.sav, 3]) 


```


```{r plot, message=FALSE}

labels <- month.abb[1:12]

# combine data from current and previous years

comb.dta <- rbind(total.dta.long, prev.total.dta.long)

comb.dta %>% mutate(Year = factor(Year)) %>%
  ggplot(aes(x = Month, y = Amount, color = Type, linetype = Year)) +
  geom_point() + geom_line() +
          scale_x_continuous(
            breaks = 1:12,
            minor_breaks = NULL,
            label =  month.abb
          ) + 
  scale_y_continuous(label = scales::label_dollar(prefix="£")) +
  scale_linetype_manual(values = c("twodash", "solid")) +
  scale_colour_manual(values = c("red", "green3")) +
  theme_light() 


```

```{r fees}
stripe <-  read_xlsx(fn, sheet = "Stripe")
stripe_fees <- sum(stripe[, "Fee"], na.rm = TRUE)
```

## Income and Expenditure

- Total income since 1 January 2025 is £`r total.income`

- Total expenditure since 1 January 2025 is £`r total.expenditure`

- **The current year's `r ifelse(total.income - total.expenditure > 0, "surplus", "deficit")` is £`r total.income - total.expenditure`**.

- Current account balance £`r as.numeric(dta[nrows,6]) `

- Savings account balance £`r savings.balance`


\pagebreak



```{r summary, eval=TRUE}
sum_tab <- bind_rows(sum_tab, savings.sum_tab)

sum_tab %>% kbl(format.args = list(big.mark = ",")) %>%
  kable_classic(full_width = FALSE) %>%
    row_spec(0, bold = TRUE) # %>%
#  kable_styling(latex_options = "hold_position")


```



```{r exhibitions, eval=TRUE}

calc_totals <- function(tabdata){
  tot_row <- tibble(Category = "TOTAL", Income = sum(tabdata[, "Income"]), 
                    Expenditure = sum(tabdata[, "Expenditure"]))
  bind_rows(tabdata, tot_row)
}

calc_exhibitions <- function(dta, exhibition = "Open"){
  dta %>%
          mutate(Date = lubridate::ymd(Date),
                 Month = lubridate::month(Date)) %>%
        filter(Exhibition == exhibition) %>%
        group_by(Category) %>%
        summarise(
          Income = sum(Income, na.rm = TRUE),
          Expenditure = sum(Expenditure, na.rm = TRUE)
        ) %>%
    calc_totals(.)
}
 
members_dta <- calc_exhibitions(dta, "Members")


open_dta <- calc_exhibitions(dta)

```



## Members exhibition


```{r members, eval=TRUE}

nrow_mbs <- dim(members_dta)[1]

members_dta %>% kbl(format.args = list(big.mark = ",")) %>%
  kable_classic(full_width = FALSE) %>%
  row_spec(nrow_mbs, bold = TRUE, color = "white", background = "black")%>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(latex_options = "hold_position")

```


\pagebreak



```{r open, eval=FALSE}
nrow_open <- dim(open_dta)[1]
open_dta %>% kbl(format.args = list(big.mark = ",")) %>%
  kable_classic(full_width = FALSE) %>%
  row_spec(nrow_open, bold = TRUE, color = "white", background = "black") %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(latex_options = "hold_position")
  
```

*Notes*

- Stripe fees: £`r stripe_fees`.