---
title: "OAS Treasurer's Report"
author: "David Barron"
date: last-modified
date-format: DD MMMM YYYY
format:
  html:
    backgroundcolor: '#FFF8DC'
    fontcolor: '#000080'
    mainfont: Georgia
echo: false
---

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(kableExtra)
library(here)
library(readxl)

year <- 2025 # for filename
prevYear <- 2024

```

```{r}
#| label: numberformat
knit_print.integer = function(x, ...) {
  prettyNum(x, big.mark = ",")
}
registerS3method(
  "knit_print",
  "integer",
  knit_print.integer,
  envir = asNamespace("knitr")
)

# Custom print method for large numerics
knit_print.numeric = function(x, ...) {
  prettyNum(x, big.mark = ",")
}
registerS3method(
  "knit_print",
  "numeric",
  knit_print.numeric,
  envir = asNamespace("knitr")
)
```

```{r}
#| label: functions
#| warning: false
#| echo: false

# Reads data from spreadsheet containing bank transactions
read_dta <- function(fn, sheet) {
  read_xlsx(fn, sheet = sheet) |>
    filter(Category != "Transfer Out") |> # remove transfers between accounts
    filter(Category != "Transfer In")
}


# Summary of income and expenditure by month

sum_dta <- function(dta) {
  dta %>%
    mutate(Date = lubridate::ymd(Date), Month = lubridate::month(Date)) %>%
    group_by(Month) %>%
    summarise(
      Expend = sum(Expenditure, na.rm = TRUE),
      Inc = sum(Income, na.rm = TRUE)
    )
}

# calculate cummulative amounts for plotting
calc.cumsum <- function(dta) {
  dta %>%
    mutate(
      Income = cumsum(Inc),
      Expenditure = cumsum(Expend)
    ) %>%
    select(Month, Income, Expenditure)
}


# Transform data for plotting

calc.long <- function(dta) {
  dta %>%
    pivot_longer(
      cols = Income:Expenditure,
      names_to = "Type",
      values_to = "Amount"
    )
}

# Add together amounts from current and savings accounts
# Have to take account of times when there have been no changes to savings account
#  during the current month

calc.sum <- function(dta) {
  dta %>%
    group_by(Category) %>%
    summarise(
      Income = sum(Income, na.rm = TRUE),
      Expenditure = sum(Expenditure, na.rm = TRUE)
    )
}

combine_data <- function(current, savings) {
  full_join(current, savings, by = "Month") %>%
    mutate(Inc = Inc.x + Inc.y, Expend = Expend.x + Expend.y) %>%
    select(Month, Inc, Expend) %>%
    replace(is.na(.), 0) # if there have been no transactions in a month
}

```

```{r}
#| label: dataprocessing
#| echo: false
#| message: false
fn <- paste0("OAS ", year, ".xlsx")

dta <- read_dta(fn, sheet = "Current account")
savings.dta <- read_dta(fn, sheet = "Savings account")
nrows <- nrow(dta)

dta.sum <- sum_dta(dta)
savings.dta.sum <- sum_dta(savings.dta)

# Combine current and savings accounts
comb.dta <- combine_data(dta.sum, savings.dta.sum)

# Calculate cummulative income and expenditure over the year to date
dta.cumsum <- calc.cumsum(comb.dta)

num.months <- dim(dta.cumsum)[1]

# Long format for printing
dta.long <- calc.long(dta.cumsum)

total.dta.long <- dta.long %>%
  mutate(Year = year)

nrow.savings <- nrow(savings.dta)
savings.balance <- as.numeric(savings.dta[nrow.savings, 6])

sum_tab <- calc.sum(dta)

savings.sum_tab <- calc.sum(savings.dta)

total.income <- sum(sum_tab$Income)

total.expenditure <- sum(sum_tab$Expenditure)

```

```{r}
#| label: prevdata
#| message: false
#| warning: false
#| echo: false

# Read and process previous year's data
prevfn <- paste0("OAS ", prevYear, ".xlsx")

prev.dta <- read_dta(prevfn, sheet = "Current account")
prev.savings.dta <- read_dta(prevfn, sheet = "Savings account")

prev.dta.sum <- sum_dta(prev.dta)
prev.savings.dta.sum <- sum_dta(prev.savings.dta)

prev.nrows <- nrow(prev.dta)

prev.comb.dta <- combine_data(prev.dta.sum, prev.savings.dta.sum)

prev.dta.cumsum <- calc.cumsum(prev.comb.dta)

prev.dta.long <- calc.long(prev.dta.cumsum)

prev.total.dta.long <- prev.dta.long %>%
  mutate(Year = prevYear)

# Summary of income and expenditure by month

#! prev.dta.sum <- sum_dta(prev.dta)
#! prev.savings.dta.sum <- sum_dta(prev.savings.dta)

# calculate cummulative amounts for plotting

#prev.dta.cumsum <- calc.cumsum(prev.dta.sum)
#prev.savings.dta.cumsum <- calc.cumsum(prev.savings.dta.sum)

#prev.num.months <- dim(prev.dta.cumsum)[1]
#prev.num.months.sav <- dim(prev.savings.dta.cumsum)[1]

# Transform data for plotting

#prev.dta.long <- calc.long(prev.dta.cumsum)
#prev.savings.dta.long <- calc.long(prev.savings.dta.cumsum)

# Add together amounts from current and savings accounts
# Have to take account of times when there have been no changes to savings account
#  during the current month

#prev.total.dta.long.dta <- full_join(
#  prev.dta.long,
#  prev.savings.dta.long,
#  by = c("Month", "Type")
#)
#prev.total.dta.long.dta[is.na(prev.total.dta.long.dta[, 4]), 4] <- 0

#prev.total.dta.long <- prev.total.dta.long.dta %>%
#  mutate(Amount = Amount.x + Amount.y, Year = prevYear) %>% # need a Year variable for when data is combined
#  filter(!is.na(Month)) %>%
#  select(Month, Type, Amount, Year)

prev.nrow.savings <- nrow(prev.savings.dta)
prev.savings.balance <- as.numeric(prev.savings.dta[prev.nrow.savings, 6])

prev.sum_tab <- calc.sum(prev.dta)

prev.savings.sum_tab <- calc.sum(prev.savings.dta)

prev.total.income <- as.numeric(prev.dta.cumsum[12, 2])

prev.total.expenditure <- as.numeric(prev.dta.cumsum[12, 3])

```

```{r}
#| label: plot
#| echo: false
#| message: false

labels <- month.abb[1:12]

# combine data from current and previous years

comb.dta <- rbind(total.dta.long, prev.total.dta.long)

comb.dta %>%
  mutate(Year = factor(Year)) %>%
  ggplot(aes(x = Month, y = Amount, color = Type, linetype = Year)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(
    breaks = 1:12,
    minor_breaks = NULL,
    label = month.abb
  ) +
  scale_y_continuous(label = scales::label_dollar(prefix = "£")) +
  scale_linetype_manual(values = c("twodash", "solid")) +
  scale_colour_manual(values = c("red", "green3")) +
  ggtitle("Cummulative income and expenditure") +
  theme_light() +
  theme(
    panel.background = element_rect(fill = "#FFF8DC"),
    plot.background = element_rect(fill = "#FFF8DC"),
    legend.background = element_rect(fill = "#FFF8DC")
  )

```

```{r}
#| label: fees
#| echo: false

stripe <- read_xlsx(fn, sheet = "Stripe")
stripe_fees <- sum(stripe[, "Fee"], na.rm = TRUE)

sumup <- read_xlsx(fn, sheet = "SumUp")
sumup_fees <- sum(sumup[, "Fee"], na.rm = TRUE)
```

## Income and Expenditure

-   Total income since 1 January 2025 is £`r total.income`
-   Total expenditure since 1 January 2025 is £`r total.expenditure`
-   **The current year's `r ifelse(total.income - total.expenditure > 0, "surplus", "deficit")` is £`r total.income - total.expenditure`**
-   Current account balance £`r as.numeric(dta[nrows,6])`
-   Savings account balance £`r savings.balance`

```{r}
#| label: summary
#| echo: false
#| output: asis

sum_tab <- bind_rows(sum_tab, savings.sum_tab)
nrow_sum <- dim(sum_tab)[1]

sum_tab %>%
  kbl(format.args = list(big.mark = ",")) %>%
  kable_classic(full_width = FALSE, html_font = "Georgia") %>%
  # row_spec(nrow_sum, bold = TRUE, color = "white", background = "black")%>%
  row_spec(0, bold = TRUE) %>%
  row_spec(0:nrow_sum, color = "#000080")
#  kable_styling(latex_options = "hold_position")

```

```{r}
#| label: exhibitionfunctions
#| echo: false

calc_totals <- function(tabdata) {
  tot_row <- tibble(
    Category = "TOTAL",
    Income = sum(tabdata[, "Income"]),
    Expenditure = sum(tabdata[, "Expenditure"])
  )
  bind_rows(tabdata, tot_row)
}

calc_exhibitions <- function(dta, exhibition = "Open") {
  dta %>%
    mutate(Date = lubridate::ymd(Date), Month = lubridate::month(Date)) %>%
    filter(Exhibition == exhibition) %>%
    group_by(Category) %>%
    summarise(
      Income = sum(Income, na.rm = TRUE),
      Expenditure = sum(Expenditure, na.rm = TRUE)
    ) %>%
    calc_totals(.)
}


exhibition_table <- function(dta) {
  n <- dim(dta)[1]

  dta %>%
    kbl(format.args = list(big.mark = ",")) %>%
    kable_classic(full_width = FALSE, html_font = "Georgia") %>%
    row_spec(0:n, color = "#000080") %>%
    row_spec(n, bold = TRUE, color = "white", background = "black") %>%
    row_spec(0, bold = TRUE) %>%
    kable_styling(latex_options = c("hold_position"), position = "center")
}

exhibition_outcome <- function(dta) {
  n <- dim(dta)[1]
  as.numeric(dta[n, 2] - dta[n, 3])
}

```

```{r}
#| label: exhibitionprocessing
#| echo: false
#| message: false

members_dta <- calc_exhibitions(dta, "Members")

open_dta <- calc_exhibitions(dta)

ya_dta <- calc_exhibitions(dta, "Young artists")


```

## Members exhibition

```{r}
#| label: members
#| echo: false
#| output: asis

exhibition_table(members_dta)
members_outcome <- exhibition_outcome(members_dta)

```

-   **Members exhibition cost: £`r -members_outcome`**

## Young artists exhibition

```{r}
#| label: ya
#| echo: false
#| output: asis

exhibition_table(ya_dta)
ya_outcome <- exhibition_outcome(ya_dta)

```

-   **Young artists exhibition cost: £`r -ya_outcome`**

## Open exhibition

```{r}
#| label: open
#| echo: false
#| output: asis

exhibition_table(open_dta)
open_outcome <- exhibition_outcome(open_dta)

```

***Notes:***

-   The "Other income" of £850 arose because someone accidentally paid twice for a painting. It is balanced by the "Refund" for the same amount. 

-   Awaiting invoices from one exhibition organiser and St John's College.  I expect the St Johns invoice to be around £2k.
  
-   One artist is yet to pay the £30 commission that is due to us from a direct sale.

-   **Open artists exhibition surplus to date: £`r open_outcome`**

*Notes*

-   Stripe fees: £`r stripe_fees`
-   SumUp fees: £`r sumup_fees`